\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-11-20 AJB_Res2.tex
Updates to manuscript
Rewrite of allelic series section, same conclusions but rewritten. I
think we disagree with the reviewer on this point.

Page 15, line 14, discussion of Studer et al. (2012).  Studer et al. showed that there is an allelic series of tb1 in teosinte, and that this series is correlated with taxonomy of maize subspecies.  They also showed that differences between the three classes of tb1 alleles were significantly correlated with mean internode length of the uppermost lateral branch as well as two inflorescence traits, but that the classes did NOT correlate with variation in tillering.  Studer et al. concluded that there was evidence for allelic series only for the internode length and inflorescence characters, not for tillering.  Their explanation was NOT that tillering index was controlled by both tb1 and other loci, but that there was no evidence for variation in tb1 alleles affecting tillering.  Changes in other loci were invoked by Studer et al. (along with environmental effects on seed development) to explain why separate families differed from each other irrespective of the tb1 allele they
contained.  Again, at the end of this paragraph (page 15, lines 27-28), there is confusion over what is being explained, as the last sentence refers to a lack of correlation between Hopscotch genotype and tillering, but the discussion of the Studer et al. (2012) paper does not concern Hopscotch, but only an allelic series at the tb1 locus.  In fact, the Studer paper does not mention Hopscotch elements at all and the allelic series analyzed are only the coding region of the tb1 gene.
\lev{I think we should reread this section of the studer paper and check. If I remember correctly they found differences between their maize isogenic lines that were comparable to the differences they found among maize with introgressed segments of tb1}

>>>>

Page 15, line 14, discussion of Studer et al. (2012).  Studer et al. showed that there is an allelic series of tb1 in teosinte, and that this series is correlated with taxonomy of maize subspecies.  They also showed that differences between the three classes of tb1 alleles were significantly correlated with mean internode length of the uppermost lateral branch as well as two inflorescence traits, but that the classes did NOT correlate with variation in tillering.  Studer et al. concluded that there was evidence for allelic series only for the internode length and inflorescence characters, not for tillering.  Their explanation was NOT that tillering index was controlled by both tb1 and other loci, but that there was no evidence for variation in tb1 alleles affecting tillering.  Changes in other loci were invoked by Studer et al. (along with environmental effects on seed development) to explain why separate families differed from each other irrespective of the tb1 allele they
contained.  Again, at the end of this paragraph (page 15, lines 27-28), there is confusion over what is being explained, as the last sentence refers to a lack of correlation between Hopscotch genotype and tillering, but the discussion of the Studer et al. (2012) paper does not concern Hopscotch, but only an allelic series at the tb1 locus.  In fact, the Studer paper does not mention Hopscotch elements at all and the allelic series analyzed are only the coding region of the tb1 gene.
\lev{I think we should reread this section of the studer paper and check. If I remember correctly they found differences between their maize isogenic lines that were comparable to the differences they found among maize with introgressed segments of tb1}
\lev{I have reread his thesis a bunch since getting the review back, and either i am misunderstanding or the reviewer is. He is correct that the chapter/paper does not explicitly mention the Hopscotch, but if I am reading the figure correctly all of tb1 ends up in one or more of the introgressed segments from the long arm of chr1. The methods also discuss amplifying both the coding region and the control region. I don't think we are misinterpreting. The paper explicitly says that they find evidence that "different teosinte tb1 chromosomal segments possess distinct genetic factors that confer different amounts of tillering." It goes on to say that they found significant variation in tillering in their control line W22 and their conclusions for tb1 chromosomal segments with respect to tillering is "These results suggest that there are factors other than tb1 affecting phenotype. Therefore we cannot conclude that the observed differences in tillering among the teosinte introgressions are caused by tb1 alone".}

\lev{What do you guys think about changing this or leaving it as is?}

### 2014-11-20 manuscript.tex
Updates to manuscript
Rewrite of allelic series section, same conclusions but rewritten. I
think we disagree with the reviewer on this point.

A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus. \citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated their phenotypic effects. Phenotypes were shown to cluster by taxon, indicating \emph{tb1} may underlie morphological diversification of \emph{Zea}. Additional analysis in \citet{StuderDoebley2012} suggested tillering index was controlled both by \emph{tb1} and loci elsewhere in the genome. Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}). Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{te1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{te1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}. \emph{tru1}, for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number; \citealt{Li2012}). Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.

>>>>

A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus. \citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated their phenotypic effects. 
Their findings suggest that different teosinte \emph{tb1} segments contain distinct genetic factors that differentially contribute to tillering in teosinte. Furthermore, \citet{StuderDoebley2012} found significant variation in tillering within their control maize lines, concluding that variation in \emph{tb1} alone cannot account for the differences in tillering, and that other genetic factors must also be contributing to differences in tillering in teosinte. 
Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}). Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{te1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{te1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}. \emph{tru1}, for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number; \citealt{Li2012}). Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
